I/O simulation model with Linux page cache, integration and evaluation in SimGrid framework

### 2020-12-07 paper_io_sim.tex
Comments in III.D

        Our experiments compared real executions with the original WRENCH,
        our WRENCH-cache extension, and our prototype Python simulator. We
        evaluated our page cache simulation in single-threaded and
        multi-threaded applications accessing data on local and network
        file systems.

            We used a pipeline of three sequential tasks where each task reads the output of the previous task,
            increments every byte of the input to emulate real
            processing, and writes the resulting data to disk. The first file was provided as input. We also used
            a real application pipeline \textcolor{red}{[Maybe briefly
            describe this application's purpose (e.g. a standard fMRI
            preprocessing piepeline)]}, to evaluate the applicability of
            our simulation model.

            We use a dedicated cluster hosted at Concordia University to conduct
            the experiments. The cluster consists of one login node, 9 compute nodes
            and 4 storage nodes connected with a network bandwidth of 25 Gbps.
            Each compute node has two 16 cores Intel(R) Xeon(R) Gold 6130 CPU @ 2.10GHz,
            275~GB (256~GiB) of RAM and 6 SSDs, 450~GB each with XFS file system,
            378~GB of tmpfs and 126~GB of devtmpfs file system.
            The cluster is run on CentOS 8.1 with NFS version 4.
            We use \textbf{\textit{atop}} and \textbf{\textit{collectl}} as tools to
            monitor and collect memory status and disk throughput
            on the cluster.

            In simulated platform description, because the disk bandwidths in SimGrid
            are symmetrical, we measure the read/write bandwidths on our cluster and
            take the average bandwidths.
            The benchmark results of bandwidths and the values used in the simulators
            in our all experiments are shown in Table~\ref{table:benchmark}.

            \begin{table}[htbp]
            \centering
            \begin{tabularx}{\columnwidth}{|l
            |>{\centering\arraybackslash}X
            |>{\centering\arraybackslash}X
            |>{\centering\arraybackslash}X|}
            \hline
                Bandwidths (MBps)  & Cluster (real) & Python simulator & WRENCH simulator\\
            \hline
                Memory read  & 6860	& 4812	 & 4812\\
                Memory write & 2764	& 4812 & 4812\\
                Local disk read & 510 & 465 & 465\\
                Local disk write & 420 & 465	 & 465\\
                Remote disk read & 515 & - & 445\\
                Remote disk write & 375 & - & 445\\
                Network bandwidth & 3000 & - & 3000\\
            \hline
            \end{tabularx}
            \caption{Bandwidth benchmark results and simulator configurations.
            The bandwidths used in simulators are average of read and write bandwidths
            measured on the cluster.
            Remote disk and network are not simulated in Python simulator.}
            \label{table:benchmark}
            \end{table}

            As the pipelines are single-core, we create simulated tasks,
            each of them runs on only one core.
            While in Python, computation time can be directly configured,
            this time is injected to the WRENCH simulators with number of flops,
            which is calculated with CPU speed and compute duration.
            In our experiments, the CPU speed is set to 1 Gflops.
            Finally, because WRENCH simulates applications base on network-communication,
            we use an infinite bandwidth to eliminate network latency for local I/Os.
            The details of simulated platform and the source code of the simulator
            used in our experiments are available in this commit:
            \url{https://github.com/wrench-project/wrench/tree/ec6b43561b95977002258c0fe37a4ecad8f1d33f/examples/basic-examples/io-pagecache}.

            \subsubsection{Single-threaded evaluation}

            The first experiment is single-threaded, where memory bandwidth
            and disk bandwidth are dedicated.
            Our goal is to compare the detailed simulation results of each task
            in the pipeline including read/write time, memory behavior and
            then validate the correctness of the simulation model.

            The pipeline used in this experiment consists of three tasks,
            each task reads a file, increases every byte and writes output file
            to local disk.
            We use the input sizes of 20~GB, 50~GB, 75~GB, 100~GB and run
            the pipeline on the cluster.
            The pipeline is simulated with the Python prototype simulator,
            original WRENCH and WRENCH-cache simulator.

            \subsubsection{Multi-threaded evaluation}

            In the second scenario, we evaluate the simulator in a concurrent I/O, 
            bandwidth-sharing scenario with a multi-threaded application.             
            We vary the number of concurrent pipelines running on a multi-core host.

            All the task in the pipelines read input files from and write output files 
            to a local disk. 
            As a compute node of the cluster has 32 cores CPU and 256~GB of RAM,  
            we create 32 input files with the size of 3~GB each and vary the number of 
            concurrent pipelines from 1 to 32.

            Because implementation of a multi-threaded simulator in Python is 
            expensive, and we have our model implemented in WRENCH, 
            we only compare the results of simulators with original WRENCH, 
            WRENCH-cache and real pipelines in this experiment. 
            The results are compared in total makespan of the pipelines, 
            cumulative read time and cumulative write time.

            \subsubsection{Evaluation on NFS}

            Since network file systems are very common in HPC, we conduct 
            the next experiment with the same scenario as in the multi-threaded experiment, 
            except that files are stored on a node, read and written by another node. 
            The compute node is the node used in the multi-threaded experiment, 
            and the capacity of the disk used on the storage node is 400 GB.

            In NFS, client cache and server cache are enabled in reading, but only 
            server write-through cache is involved in writing. 
            Although write-through does not have any impacts writing performance, 
            caching data on server side allows server to avoid disk reads, which are 
            significantly longer than cache reads.
            Thus, we simulate page cache for reading on both client and server node.
            In writing, write-though is simulated with data being written directly to 
            server storage and then added to server cache.

            \subsubsection{Simulation of a real application}

            Finally, we consider a real pipeline from the neuroimaging domain.             
            \textcolor{red}{[Description of the real pipeline with nighres 
            (including the workflow engine if it's Dask)]}.

            We simulate this pipeline with simulators using original WRENCH and 
            WRENCH-cache. 
            The simulation results are then compared with the real results. 
            Because our work focuses on I/O time, we assume that CPU time is 
            correctly modeled and use the CPU time measured in the real pipelines 
            to setup our simulated workflow. The results are compared in terms of 
            read/write time and total makespan.

>>>>

        Our experiments compare real executions with the original WRENCH,
        with our WRENCH-cache extension, and with our prototype Python simulator. We
        evaluate our page cache simulation in single-threaded and
        multi-threaded applications, accessing data on local and remote
        file systems.

        We simulate an application with three single-core, sequential tasks
        where each task reads the file produced by the previous task,
        increments every byte of this file to emulate real processing, and
        writes the resulting data to disk. We also use a real application
        \textcolor{red}{[Maybe briefly describe this application's purpose
        (e.g. a standard fMRI preprocessing piepeline)]}, to evaluate the
        applicability of our simulation model.

        The real application runs on a dedicated cluster hosted at
        Concordia University, with one login node, 9 compute nodes, and 4
        storage nodes connected with a network bandwidth of 25 Gbps. Each
        compute node has 2 $\times$ 16-core Intel(R) Xeon(R) Gold 6130 CPU
        @ 2.10GHz, 275~GB (256~GiB) \tristan{are you sure it's not 256GB?
        pls check.} of RAM, 6 $\times$ SSDs of 450~GB each with the XFS
        file system, 378~GB of tmpfs, and 126~GB of devtmpfs file system.
        Nodes run CentOS~8.1 with NFS version 4. We use the \texttt{atop}
        and \texttt{collectl} tools to monitor and collect memory status
        and disk throughput.

        To parametrize our simulators, we benchmarked the bandwidths of
        memory, local disk, remote disk (NFS), and network
        (Table~\ref{table:benchmark}). Since SimGrid only supports
        symmetrical bandwidths, we use the mean of the read and write
        bandwidth values in our experiments.
            \begin{table}[htbp]
            \centering
            \begin{tabularx}{\columnwidth}{|c
            |>{\centering\arraybackslash}X
            |>{\centering\arraybackslash}X
            |>{\centering\arraybackslash}X|}
            \hline
                Device  & Cluster (real) & Python simulator & WRENCH simulator\\
            \hline
                Memory read  & 6860	& 4812	 & 4812\\
                Memory write & 2764	& 4812 & 4812\\
                Local disk read & 510 & 465 & 465\\
                Local disk write & 420 & 465	 & 465\\
                Remote disk read & 515 & - & 445\\
                Remote disk write & 375 & - & 445\\
                Network bandwidth & 3000 & - & 3000\\
            \hline
            \end{tabularx}
            \caption{Bandwidth benchmarks (MBps) and simulator configurations.
            The bandwidths used in the simulations are the average of the read and write bandwidths
            measured in the cluster.
            Remote disk and network accesses are not simulated in the Python simulator. \tristan{table could be prettified}}
            \label{table:benchmark}
            \end{table}

            As our focus is on I/O rather than compute, we measured
            application task durations on a cluster node and used these
            durations in our simulation. For the Python simulator,
            execution times were directly used in the simulation. For
            WRENCH and WRENCH-cache, we determined the corresponding number
            of flops on a 1~Gflops CPU and used these values in the
            simulation (see Table~\ref{table:tasks} \tristan{add a table to describe execution times}). The final simulated
            platform and application used in our experiment are avaiable at
            \url{https://github.com/wrench-project/wrench/tree/ec6b43561b95977002258c0fe37a4ecad8f1d33f/examples/basic-examples/io-pagecache}.

            % Finally, because WRENCH simulates applications base on network-communication,
            % we use an infinite bandwidth to eliminate network latency for local I/Os.

            Our first experiment simply runs one instance of the
            application on a single cluster node, with different input file
            sizes (20~GB, 50~GB, 75~GB, 100~GB), and with all I/Os directed
            to the same local disk.

            Our second experiment runs concurrent instances of the
            application on a single node, all application instances
            operating on different files of size 3~GB stored in the same
            local disk, sharing the disk bandwidth. We vary the number of
            concurrent application instances from 1 to 32 since cluster
            nodes have 32 CPU cores.

            Our third experiment occurs in the same configuration as the
            previous one, albeit on a 400-GB NFS-mounted remote disk
            mounted from another compute node. \tristan{As is commonly
            configured in HPC environments to avoid data loss,?} NFS client
            and server read caches are enabled, but server writes are configured with
            write-through cache instead of write back \tristan{how about client write
            cache?}. Therefore, all the writes happen at disk bandwidth, but
            reads can benefit from cache hits. We simulate write-through by
            writing data directly to disk and adding them to the server
            cache \tristan{it it part of the IOController? it would deserve
            a few sentences in the methods.}.

            Finally, we simulate a real application from the neuroimaging
            domain. \textcolor{red}{[Description of the real pipeline with
            nighres (including the workflow engine if it's Dask)]}.
            We simulate this pipeline with simulators using original WRENCH
            and WRENCH-cache. Because our work focuses on I/O time, we
            assume that CPU time is correctly modeled and use the CPU time
            measured in the real pipelines to setup our simulated
            application.
